ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python

### 2016-08-23 doc/paper.tex
example in paper
typo
update paper

\usepackage{fancyhdr} \usepackage{hyperref}

@@

ScPy allows users to overcome these limitations. By embedding the Python programming language
within SuperCollider, it is possible to do numerical processing operations which would be too slow
to perform in real time with SuperCollider alone. Through access to the NumPy library, previously
difficult or impossible phase vocoder operations become trivial.  Although our usage of this
library for testing purposes has mostly focused on FFT based operations, support exists for passing
arbitrary data buffers to Python, which can be used to process numerical data for many other
purposes as well.

@@

\section{Literature Review}

@@

>>>>

\usepackage{fancyhdr} \usepackage{hyperref} \usepackage{listings}

@@

ScPy enables users to overcome these limitations. By embedding the Python programming language
within SuperCollider, it is possible to do numerical processing operations which would be too slow
to perform in real time with SuperCollider alone. Through access to the NumPy library, previously
difficult or impossible phase vocoder operations become trivial. Although our usage of this library
for testing purposes has mostly focused on FFT based operations, support exists for passing
arbitrary data buffers to Python, which can be used to process numerical data for many other
purposes as well.

\section{Methodology}

During the initial planning stage of the project, we wanted to essentially provide a set of
fundamental operations on spectral data which could be composed together to perform any conceivable
effect. It was also a requirement that these operations would be syntactically concise in order to
enable easy and fast experimentation. Ideally, the system would be able to handle complex effects
in real-time.

Although designing a domain-specific language (DSL) for this task was briefly considered, we
eventually decided to use Python for a number of reasons. First, Python includes an excellent C API
which makes embedding it within other languages trivial in comparison to many other options.
Second, Python's syntax is easily readable and concise -- it clearly adheres to our goals. Finally,
Python has a massive number of community built libraries for performant scientific and numerical
computing. With a few small exceptions, these libraries contained every one of the fundamental
spectral operations we could think of.

@@

@@

\lstset{ caption={SuperCollider boilerplate for no-op FFT effect with ScPy.} }
\lstinputlisting[ basicstyle=\footnotesize\ttfamily]{../examples/template.sc}

ScPy is written mostly in C++, but with a small SuperCollider class library to connect the C++
back-end, and a small Python library that provides some useful operations. Due to differences
between these languages, connecting them was somewhat awkward in some cases.

One issue involves the process of passing data between SuperCollider and C++. Since UGens are
designed to work primarily with data streams, adding support for passing in many types of data came
with a number of challenges. Essentially, all data must be serialized into an array of floats.
Strings must be converted to an array of ASCII character codes, and prepended with their length.
Arguments passed through to the Python code can have one of many types, and therefore type
information must be encoded as well. However, just passing the type of a variable is not enough to
know what we can do with it, since we also want to know if it inherits from a class we can use.
